Massively Parallel Pseudo-Spectral DNS and~LES for Particle-Laden Turbulent Flows under Two-Way Momentum Coupling

### 2023-01-21 msc-thesis.tex
updated with reviewed chapter 2 text (+ figure and table excerpts)

\streszczenie{
Przepływy turbulentne z cząstkami fazy rozproszonej zachodzą nieustannie w środowisku naturalnym oraz znajdują liczne zastosowania w procesach przemysłowych.
Ważnym tego przykładem są chmury atmosferyczne, które składają się z ogromnej ilości małych kropel wody i kryształków lodu.
Wyjaśnienie wzajemnego oddziaływania powietrza (płynu) i aerozolu (cząstek) jest istotne w kontekście poznania mechanizmów rządzących powstawaniem i przebiegiem zjawisk atmosferycznych.
Wiedza w tym zakresie jest użyteczna w obszarze meteorologii i służy do opracowywania coraz to bardziej realistycznych modeli powstawania opadu z kropel chmurowych.
To z kolei prowadzi do podniesienia dokładności i sprawdzalności numerycznych prognoz pogody.
Jednym ze sposobów badania przepływów dyspersyjnych są symulacje numeryczne. 
\newline \indent
Powszechnie stosowanym do tego celu narzędziem, zapewniającym wysoką dokładność, są tzw. bezpośrednie symulacje numeryczne (direct numerical simulation, DNS). 
Za pomocą symulacji DNS można pozyskiwać statystyki zderzeniowe cząstek, które to są użyteczne jako parametry w modelach obejmujących większe skale przestrzenne.
Symulacje DNS są~jednak kosztowne obliczeniowo, co ogranicza ich stosowalność do siatek o średnich rozmiarach (${\sim 2048^3}$ węzłów -- przy współcześnie dostępnych zasobach obliczeniowych).
DNS nie pozwalają modelować turbulencji charakteryzowanej wyższymi liczbami Reynoldsa ($R_{\lambda} \sim 10^4$), czyli takiej która obserwowana jest w powietrzu atmosferycznym. \newline \indent
Metoda dużych wirów (large-eddy simulation, LES) stanowi alternatywę dla DNS.
W tej metodzie drobnoskalowa turbulencja nie jest bezpośrednio rozwiązywana, ale jest parametryzowana przy użyciu modelu podskalowego.
Umożliwia to modelowanie przepływów turbulentnych o większej liczbie Reynoldsa przy użyciu siatek o mniejszej liczbie węzłów.
Ograniczenie złożoności obliczeniowej odbywa się jednak kosztem mniejszej precyzji i fizycznej dokładności.
Niniejsza praca stanowi porównanie obu metod (DNS i LES), zarówno w kwestii fizycznej dokładności, jak i wydajności, skupiając się w szczególności na symulacjach z dwustronnym sprzężeniem pędu pomiędzy płynem i cząstkami. 
Zostało wykazane, że LES stanowi obiecującą odpowiedź na ograniczenia DNSu, pomimo pewnych niedokładności wynikających z filtrowania turbulencji w najmniejszych skalach.
Różnice te, w większości przypadków, stają się mniejsze, gdy uwzględni się dwustronne sprzężenie pędu, w szczególności dla cząstek poddanych działaniu grawitacji.
\newline \indent
Przeprowadzone zostały również analizy kosztów obliczeniowych obu metod.
Choć zastosowanie LES zapewnia ogromną oszczędność w obliczeniach związanych z turbulentnym przepływem płynu, to jednak koszt ten znacząco rośnie wraz z liczbą cząstek.
Niezbędna ilość symulowanych cząstek może zaś być duża, gdyż efekty dwustronnego sprzężenia pędu objawiają się wyraźniej w układach z względnie wysokim stosunkiem masy cząstek i płynu (pomiędzy $0.1$~i~$1$).
Obliczenia odnoszące się do dynamiki cząstek stanowią więc bardziej istotne ograniczenie dla LESu niż dla DNSu, gdzie złożoność obliczeń związanych z~płynem jest czynnikiem dominującym.
Ponadto zidentyfikowano i przeanalizowano szereg parametrów, które pośrednio wpływają na czas trwania symulacji, takie jak np. promienie kropel oraz~działanie grawitacji.
Na~koniec przedstawiono także wpływ preferencyjnego koncentrowania się cząstek na wariancję ich rozkładu pomiędzy równolegle wykonywanymi procesami.
Ten czynnik ma~również istotne znaczenie w kontekście całkowitej wydajności symulacji.
}

\renewcommand{\abstractname}{Abstract}
\begin{abstract}
Turbulent fluid flows with dispersed particles are a common occurrence in nature, as~well~as in many technological processes.
An important example of such systems are atmospheric clouds that consist of a huge number of small water droplets and ice crystals suspended in air.
The~understanding of interplay between air (fluid) and aerosol (particles) is crucial for increasing the accuracy and reliability of numerical weather forecasts through more realistic modelling of atmospheric phenomena.
Performing numerical simulations on a fine scale is one way of studying such processes.

Direct numerical simulations (DNS) prove to be a relatively precise tool that is widely used in these studies.
DNS can be used to obtain the particle collision statistics obtained from these simulations may be further utilised as parameterisations for larger-scale models.
These simulations, however, are costly in terms of computational complexity, hence their application is limited to moderate-sized grids ($\sim 2048^3$ nodes, with the computing power of modern supercomputers).
For that reason DNS are not able to resolve turbulence characterised by~higher Reynolds numbers ($R_{\lambda} \sim 10^4$), which is observed in atmospheric air.

Large-eddy simulations (LES) provide an alternative approach to DNS.
In this method small-scale turbulence is not directly resolved but parameterised using a subgrid-scale model.
This reduces the grid size required to achieve higher Reynolds numbers (and, thus, computational complexity) at the cost of diminished precision and physical fidelity. 
This thesis provides comparison of DNS and LES, both in terms of physical fidelity and performance, focusing primarily on simulations under two-way momentum coupling between the fluid and particles.
LES is shown to be a promising solution to limitations of DNS, even though it~exhibits certain inaccuracies that may be attributed to the filtering of small-scale vortical structures.
These discrepancies, for the most part, become less pronounced when two-way momentum coupling is considered, especially for settling particles.

Furthermore, a broader study of computational costs for both methods is presented.
Even~though LES provides immense performance advantage when it comes to fluid simulation, it~is~highly susceptible to the growing number of individually tracked particles.
The amount of particles that are needed may be considerable, as effects of two-way momentum coupling manifest when particle mass loadings are relatively high (between $0.1$ and $1$).
Thus, computations related to particles are recognised to be a much more significant bottleneck for~LES than for DNS, where the complexity of fluid simulation is the largest concern.
In~addition, several parameters that indirectly influence the execution time are identified and analysed, such as droplet radii or effects of gravity.
Finally, it is shown that the preferential concentration of particles may affect the variance of their distribution between parallel processes and~consequently influence the overall simulation performance.

@@

>>>>

\streszczenie{
Przepływy turbulentne z cząstkami fazy rozproszonej zachodzą nieustannie w środowisku naturalnym oraz znajdują liczne zastosowania w procesach przemysłowych.
Ważnym tego przykładem są chmury atmosferyczne, które składają się z ogromnej ilości małych kropel wody i kryształków lodu.
Wyjaśnienie wzajemnego oddziaływania powietrza (płynu) i aerozolu (cząstek) jest istotne w kontekście poznania mechanizmów rządzących powstawaniem i przebiegiem zjawisk atmosferycznych.
Wiedza w tym zakresie jest użyteczna w obszarze meteorologii i służy do opracowywania coraz to bardziej realistycznych modeli powstawania opadu z kropel chmurowych.
To z kolei prowadzi do podniesienia dokładności i sprawdzalności numerycznych prognoz pogody.
Jednym ze sposobów badania przepływów dyspersyjnych są symulacje numeryczne. 
\newline \indent
Powszechnie stosowanym do tego celu narzędziem, zapewniającym wysoką dokładność, są tzw. bezpośrednie symulacje numeryczne (direct numerical simulation, DNS). 
Za pomocą symulacji DNS można pozyskiwać statystyki zderzeniowe cząstek, które to są użyteczne jako parametry w modelach obejmujących większe skale przestrzenne.
Symulacje DNS są jednak kosztowne obliczeniowo, co ogranicza ich stosowalność do siatek o średnich rozmiarach (${\sim 2048^3}$ węzłów -- przy współcześnie dostępnych zasobach obliczeniowych).
DNS nie pozwalają modelować turbulencji charakteryzowanej wyższymi liczbami Reynoldsa ($R_{\lambda} \sim 10^4$), czyli takiej która obserwowana jest w powietrzu atmosferycznym. \newline \indent
Metoda dużych wirów (large-eddy simulation, LES) stanowi alternatywę dla DNS.
W tej metodzie drobnoskalowa turbulencja nie jest bezpośrednio rozwiązywana, ale jest parametryzowana przy użyciu modelu podskalowego.
Umożliwia to modelowanie przepływów turbulentnych o większej liczbie Reynoldsa przy użyciu siatek o mniejszej liczbie węzłów.
Ograniczenie złożoności obliczeniowej odbywa się jednak kosztem mniejszej precyzji i fizycznej dokładności.
Niniejsza praca stanowi porównanie obu metod (DNS i LES), zarówno w kwestii fizycznej dokładności, jak i wydajności, skupiając się w szczególności na symulacjach z dwustronnym sprzężeniem pędu pomiędzy płynem i cząstkami. 
Zostało wykazane, że LES stanowi obiecującą odpowiedź na ograniczenia DNSu, pomimo pewnych niedokładności wynikających z filtrowania turbulencji w najmniejszych skalach.
Różnice te, w większości przypadków, stają się mniejsze, gdy uwzględni się dwustronne sprzężenie pędu, w szczególności dla cząstek poddanych działaniu grawitacji.
\newline \indent
Przeprowadzone zostały również analizy kosztów obliczeniowych obu metod.
Choć zastosowanie LES zapewnia ogromną oszczędność w obliczeniach związanych z turbulentnym przepływem płynu, to jednak koszt ten znacząco rośnie wraz z liczbą cząstek.
Niezbędna ilość symulowanych cząstek może zaś być duża, gdyż efekty dwustronnego sprzężenia pędu objawiają się wyraźniej w układach z względnie wysokim stosunkiem masy cząstek i płynu (pomiędzy $0.1$~i~$1$).
Obliczenia odnoszące się do dynamiki cząstek stanowią więc bardziej istotne ograniczenie dla LESu niż dla DNSu, gdzie złożoność obliczeń związanych z~płynem jest czynnikiem dominującym.
Ponadto zidentyfikowano i przeanalizowano szereg parametrów, które pośrednio wpływają na czas trwania symulacji, takie jak np. promienie kropel oraz~działanie grawitacji.
Na koniec przedstawiono także wpływ preferencyjnego koncentrowania się cząstek na wariancję ich rozkładu pomiędzy równolegle wykonywanymi procesami.
Ten czynnik ma również istotne znaczenie w kontekście całkowitej wydajności symulacji.
}

\renewcommand{\abstractname}{Abstract}
\begin{abstract}
Turbulent fluid flows with dispersed particles are a common occurrence in nature, as well as in many technological processes.
An important example of such systems are atmospheric clouds that consist of a huge number of small water droplets and ice crystals suspended in air.
The understanding of interplay between air (fluid) and aerosol (particles) is crucial for increasing the accuracy and reliability of numerical weather forecasts through more realistic modelling of atmospheric phenomena.
Performing numerical simulations on a fine scale is one way of studying such processes.

Direct numerical simulations (DNS) prove to be a relatively precise tool that is widely used in these studies.
DNS can be used to obtain the particle collision statistics obtained from these simulations may be further utilised as parameterisations for larger-scale models.
These simulations, however, are costly in terms of computational complexity, hence their application is limited to moderate-sized grids ($\sim 2048^3$ nodes, with the computing power of modern supercomputers).
For that reason DNS are not able to resolve turbulence characterised by higher Reynolds numbers ($R_{\lambda} \sim 10^4$), which is observed in atmospheric air.

Large-eddy simulations (LES) provide an alternative approach to DNS.
In this method small-scale turbulence is not directly resolved but parameterised using a subgrid-scale model.
This reduces the grid size required to achieve higher Reynolds numbers (and, thus, computational complexity) at the cost of diminished precision and physical fidelity. 
This thesis provides comparison of DNS and LES, both in terms of physical fidelity and performance, focusing primarily on simulations under two-way momentum coupling between the fluid and particles.
LES is shown to be a promising solution to limitations of DNS, even though it exhibits certain inaccuracies that may be attributed to the filtering of small-scale vortical structures.
These discrepancies, for the most part, become less pronounced when two-way momentum coupling is considered, especially for settling particles.

Furthermore, a broader study of computational costs for both methods is presented.
Even though LES provides immense performance advantage when it comes to fluid simulation, it is highly susceptible to the growing number of individually tracked particles.
The amount of particles that are needed may be considerable, as effects of two-way momentum coupling manifest when particle mass loadings are relatively high (between $0.1$ and $1$).
Thus, computations related to particles are recognised to be a much more significant bottleneck for LES than for DNS, where the complexity of fluid simulation is the largest concern.
In addition, several parameters that indirectly influence the execution time are identified and analysed, such as droplet radii or effects of gravity.
Finally, it is shown that the preferential concentration of particles may affect the variance of their distribution between parallel processes and consequently influence the overall simulation performance.

@@

This Chapter focuses on ascertaining the physical fidelity of LES results in comparison with those provided by DNS.
Since LES partially replaces direct resolution of flow with subgrid-scale parameterisation, some difference of results is expected.
This is also true for particle statistics at contact distance, as they are especially sensitive to the effects of turbulence at smaller length scales.
In this chapter, basic statistics and energy spectra of background turbulent flow shall be presented for both methods, followed by particle statistics and the effects of turbulence modulation by particles in TWC simulations.
More importantly, broader analysis of key particle collision statistics---radial distribution function, radial relative velocity, and collision kernels---is provided with results from both DNS and LES.
Simulations with both one- and two-way momentum coupling are considered, both with and without gravity.

\section{Turbulent Flow Statistics}
\label{sc:ch2.flow}

The standard procedure in simulations using the pseudo-spectral code introduced in Chapter \ref{ch:ch1} is, at first, to obtain statistically stationary fluid flow without particles and estimate its statistics as a baseline.
These statistics are discussed in two following sections.
Later, such background flow is randomly seeded with particles and the entire system is simulated until stationary state is restored.
At that point particle statistics may be collected.
Such approach not only reduces time required by particle-laden simulations to achieve desired state but also provides a point of reference for studying influence of particles on the statistics of underlying turbulence in siumulations under two-way momentum coupling..

\subsection{Basic Statistics of Turbulent Flow}
\label{ssc:ch2.flow.bstat}

When no specific boundaries are defined, the model of homogeneous and isotropic turbulence is abstracted from any physical scale. 
More precisely, without introduction of any reference objects with well defined physical dimensions (e.g. domain boundaries, or -- as in this study -- particles with well defined physical radii) turbulence parameters may be applied at arbitrary scales, from small cells of clouds to the global atmospheric circulation.
For that reason, parameters and statistics of turbulent flow presented below are expressed in nondimensional spectral (DNS) units, as the most important indicator is the range of turbulent scales being resolved and not physical size of the domain.
It should be added that the characteristic length scale results from the domain size. 
In turn, the time and velocity scales are determined by the rate at which kinetic energy is supplied by the forcing scheme.

There are three main numeric parameters that influence the resulting turbulent flow and its statistics that are, for the most part, fixed for all simulations in this study.
These include:
\begin{itemize}
\item the number of spatial grid nodes in each dimension, $N$, which primarily determines range of turbulent scales that can be directly resolved in simulations (here $N=256$ for DNS and $N=64$ for LES);
\item the numerical (molecular) kinematic viscosity of the fluid, $\nu$, is an important parameter that is required to balance out the large-scale forcing with enough energy dissipation to obtain viable energy spectrum and desired range of resolved scales of turbulence; in all simulations value of $1.5 \times 10^{-3}$ is used, as it ensures numerical stability and maximises obtained Reynolds number;
\item the time step, $\Delta t$, ensures numerical stability of the simulation.
\end{itemize}
Note that for the particle-free flow time step must be set to a value that guarantees that Courant number is lower than $0.3$, which is established as a stability and accuracy threshold for the pseudo-spectral method.
Recall, that Courant (or Courant--Friedrichs--Lewy, CFL) number is defined as follows:
$$ \text{CFL} = \max ( || \mathbf{U} ||_1 ) \frac{\Delta t}{\Delta x} , $$
where $|| \cdot ||_1$ is a 1-norm, maximum of which is taken over all nodes in the grid, and $\Delta x$ is spatial spacing, here equal to $(2 \pi) / N$. 
The default time step for most simulations was set to $9 \times 10^{4}$, but in some cases, due to complex interactions with particles, certain adjustments were necessary (for more details, see discussion in Section \ref{ssc:ch2.flow.part}).

Beyond purely quantitative parameters there are some more ''structural'' choices that implicitly affect simulations and their results.
One such important choice involves type of forcing scheme being used (here we exclusively use deterministic scheme, see Section \ref{sc:ch1.pspec}).
Even more options come with LES method, since the choice of subgrid-scale model being used is crucial for the final outcome of simulations.
As discussed in Section \ref{sc:ch1.les}, this study employs spectral-eddy viscosity model that was already used in several previous studies \parencite{Yang2008, Jin2010, Rosa2017}.
Such model allows for some degree of customisation via model parameter $C_K$ (see Equation \ref{eqn:sgs-2}) which in this study is fixed at $2.5$.
Note, that all aforementioned studies used only stochastic forcing and this one, according to author's knowledge, is the first where LES is used with deterministic forcing scheme (with pseudo-spectral solver).
For that reason, extra effort was made to validate proper choice of $C_K$ parameter as elaborated in Appendix \ref{app:sgs}.

To gain some insight into properties of turbulence, we gather some standard flow statistics that are time-averaged over some period in which statistically stationary state of the flow was attained, and volume-averaged over the entire domain.
Firstly, two basic statistics that are calculated directly from velocity field and energy spectrum of the flow are: 
\begin{itemize}
\item $u'$ -- root-mean-square (RMS) fluctuating velocity -- standard deviation of the fluid velocity field with respect to the mean flow (which is ensured to be zero); physically, it is related to the intensity of turbulence; and
\item $\epsilon$ -- energy dissipation rate -- rate at which turbulence kinetic energy is converted into thermal internal energy; it may be computed as follows:
\begin{equation}
\epsilon  = 2 \int_{k_{0}}^{k_{\max}} \nu k^{2} E(k) dk
\label{eqn:eps}
\end{equation}
where $E(k)$ denotes the turbulent kinetic energy associated with wavenumber $k$ in the spectral space, and $\nu$ is numerical kinematic viscosity.  
\end{itemize}
Furthermore, some derived statistics that are commonly used to characterise different scales of turbulence are provided. These include: $\eta$ -- the Kolmogorov length scale, $\tau_{k}$ -- the Kolmogorov time scale, $L_{s}$ -- the integral length scale, $\lambda$ -- the transverse Taylor microscale, $T_{e}$ -- the large-eddy turnover time.
Finally, two important control and target statistics (besides CFL) are computed. These include:
\begin{itemize}
\item $R_{\lambda} = u' \lambda / \nu$ -- Taylor-microscale Reynolds number -- Reynolds number associated with inertial subrange and used to measure the range of resolved turbulent scales (in this study, this value is meant to be maximised given fixed grid scale $N$); and
\item $k_{\max} \eta$ -- spatial resolution parameter that serves as an important control value in DNS; it should be minimised in order to maximise amount of resolved turbulence (i.e. $R_{\lambda}$), but its value must be greater than $1$ to ensure that fine scales of turbulence can be resolved.
\end{itemize}
Note, that for studies oriented towards maximising $R_{\lambda}$ (as this one is) it is customary to aim for $1 < k_{\max} \eta < 2$, but in other studies resolution requirements may differ \parencite{Buaria2019} as higher values of $k_{\max} \eta$ ensure that simulations are more numerically stable.
Also, in TWC simulations it is argued that larger values of $k_{\max} \eta$ may be associated with more stability in inter-phase interpolation \parencite[see][]{Rosa2020}.

Table \ref{tab:flow-stats} contains defining parameters and gathered statistics of background turbulent flows for both DNS and LES.
The measurement uncertainties for these statistics were calculated as in \textcite{Rosa2013}, accounting for the autocorrelation of flow statistics in the statistically stationary state.
The parameters introduced above are calculated in a way specific for DNS simulations, hence may not directly translate to LES flows but still serve as helpful indicators when comparing two methods (especially, $R_{\lambda}$).
There is no agreement between authors whether to present more \parencite{Rosa2017} or less \parencite{Yang2008} of such DNS statistics for LES as well.
This is mainly due to the dependence of many statistics on numerical viscosity, $\nu$, which is modelled in subgrid-scale model and depends on the wavenumber, $k$ (see Equation \ref{eqn:sgs-1}), not being constant as in DNS.
For that reason, most of statistics for LES in Table \ref{tab:flow-stats} (marked with asterisk) are listed with ``effective'' values which are calculated in a way that accounts for subgrid-scale modelling on viscosity (see Appendix \ref{app:sgs} for details).

\begin{table}[h]
\centering
\begin{tabular}{llll}
 & & & \scriptsize{\textcite{Rosa2013}}  \\
 & \textbf{DNS} & \textbf{LES*} & \textbf{DNS}  \\ \hline
$N$ & $256$ & $64$ & $256$ \\
$\Delta t \cdot 10^{4}$ & $9.0$ & $9.0$ & $9.0$  \\
$\nu \cdot 10^{3}$ & $1.5$ & $1.5$ & $1.1$  \\

$u'$ & $0.871 \pm 0.001 $ & $0.861 \pm 0.001 $ & $0.868 \pm 0.002 $ \\
$\epsilon$ & $0.212 \pm 0.002 $ & $0.200 \pm 0.001$* & $0.200 \pm 0.003 $  \\

$\eta \cdot 10^{2}$ & $1.125 \pm 0.003$ & $2.078 \pm 0.007$* & $0.903 \pm 0.037$  \\
$\tau_{k} \cdot 10^{2}$ & $8.436 \pm 0.043$ & $15.755 \pm 0.122$* & $7.420 \pm 0.620$  \\
$L_{s}$ & $1.462 \pm 0.003$ & $1.501 \pm 0.004$ & $1.496 \pm 0.005$  \\
$\lambda \cdot 10$ & $2.851 \pm 0.011$ & $5.252 \pm 0.014$* & $2.494 \pm 0.015$  \\
$T_{e}$ & $3.616 \pm 0.028$ & $3.697 \pm 0.025$* & $3.774 \pm 0.045$  \\

$k_{\max} \eta$ & $1.423 \pm 0.004 $ & --- & $1.143 \pm 0.005 $  \\
$R_{\lambda}$ & $165.91 \pm 0.51 $ & $165.01 \pm 0.11$* & $196.87 \pm 0.78 $            
\end{tabular}
\caption{Parameters and statistics, in spectral units, of background turbulent flows simulated in this study (first two columns) and from \textcite{Rosa2013} shown for results validation (third column). 
All listed simulations used deterministic forcing scheme. \\
* -- in LES, statistics marked with * are ``effective'', i.e. calculated using viscosity adjusted by subgrid-scale model; for more details see Appendix \ref{app:sgs}.}
\label{tab:flow-stats}
\end{table}

First two columns refer to particle-free flows used for all simulations in this study, DNS and LES, respectively.
The remaining column is provided as a reference and for validation of results.
DNS turbulent flow was simulated with similar parameters to the one used for numerical experiments with one-way momentum coupling in \textcite[Table 1, entry for $256^3$]{Rosa2013}, which are presented in third column of Table \ref{tab:flow-stats}.
Note that slightly smaller value of viscosity used in that study resulted in smaller resolution parameter and, consequently, $\sim 15\%$ larger Reynolds number.
Particularly interesting is the comparison between statistics for DNS and ''effective'' statistics for LES.
For ones that are directly derived from either fluid velocity field, $u'$, or its energy spectrum, $\epsilon$, there is strong agreement.
The same is true for large scale parameters, $L_s$ and $T_e$.
What is most important, despite much smaller grid size, the $R_{\lambda}$ in both cases is almost the same due to the effects of spectral-eddy viscosity of subgrid-scale model.
Also, $u'$, that is directly computed from fluid velocity field (independet from $\nu$) and associated with the magnitude of turbulence assumes very close values for both DNS and LES.  
On the other hand, we can see increase (approximately by the factor of 2) in spatial and temporal scales (both small, $\eta$ and $\tau_k$) and intermediate ($\lambda$).
No previous LES statistics are presented for comparison, as in e.g. \textcite[Table 1 therein]{Rosa2017}, since all of these studies employed stochastic forcing scheme which makes obtained statistics not directly comparable.

\subsection{Energy and Dissipation Spectra}
\label{ssc:ch2.flow.spec}

The important insight and validation for any simulation of turbulent flow is the analysis of its energy and dissipation spectra.
Since the pseudo-spectral method naturally operates in Fourier space the energy spectrum, $E(k)$, can be easily retrieved.
Such spectrum, for intermediate wavenumbers, may be compared with theoretical Kolmogorov spectrum.
More specifically, for inertial subrange, we expect obtained spectra to closely coincide with:
\begin{equation}
E_{\text{Kolm}}(k) = C \epsilon^{2/3} k^{-5/3} ,
\label{eqn:kolm}
\end{equation}
where constant $C$ was experimentally determined to be $1.5$. 
Furthermore, the energy dissipation spectrum, $D(k)$, may be easily derived for $k \in \{ k_{0}, \ldots, k_{\max} \}$ using the following formula:
$$ D(k) = 2 \nu k^{2} E(k) . $$
Energy and dissipation spectra of turbulent flows obtained by DNS and LES are shown in Figures \ref{fig:spec} and \ref{fig:diss}, respectively.
For convenience, both spectra are shown using log-log plot with proper normalisation to ensure that values are dimensionless.

\begin{figure}[h]
\centering
\includegraphics[width=8cm]{figures/2-01_spec.pdf}
\caption{
The normalised energy spectra of background turbulent flows for both DNS and LES.
Dashed line provides comparison with $k^{-5/3}$ curve of theoretical Kolmogorov spectrum (see Equation~\ref{eqn:kolm}) for inertial subrange.
}
\label{fig:spec}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=8cm]{figures/2-02_diss.pdf}
\caption{
The normalised energy dissipation spectra of background turbulent flows for both DNS and LES.
Point where $D(k)$ is expected to be maximal (for $k \eta = 0.2$) is marked by dashed line. 
}
\label{fig:diss}
\end{figure}

Figure \ref{fig:spec} clearly shows that both energy spectra follow theoretical Kolmogorov spectrum in the inertial subrange (reference line is shifted for clarity).
For LES, however, slight underestimation of the energy for intermediate wavenumbers may be observed.
This effect is expected \parencite{Jin2010} and it is attributed to the overdissipation of energy inherent in the spectral-eddy viscosity model.
Such increase of dissipation rate for wavenumbers close to filtering threshold, $k_{c} = 30$, may be better observed on Figure \ref{fig:diss}, where respective dissipation spectra are shown.
It follows the region where $D(k)$ for LES is significantly smaller due to the smaller overall energy contained in these wavenumbers, as the energy is more strongly dissipated in larger modes (for $k \eta < 0.1$) due to contributions from the spectral-eddy viscosity.
This also manifests in a shift of the maximum of $D(k)$ which occurs for smaller wavenumbers than for DNS where such peak is observed around $k \eta \approx 0.2$, which coincides with previous simulations and experimental results \parencite{Wang1993}.
Such discrepancies between DNS and LES in dissipation rates throughout inertial subrange may affect other properties of turbulent flow, as well as particle collision statistics.
Another peculiar characteristic of energy spectra is the sudden increase of energy for largest wavenumbers (close to $k_{\max}$ for DNS, or $k_{c}$ for LES).
\textcite{Rosa2017} observed similar artifacts and suggested that they are due to accumulation of energy in the high-wavenumber modes, since the transport of energy is limited there by the viscous suppression.

\subsection{Parameters of Particle-Laden Flows}
\label{ssc:ch2.flow.part}

When simulation of background turbulent flow reaches the statistically stationary state, as described in Section \ref{ssc:ch2.flow.bstat}, individually tracked particles are introduced into the system with the uniform spatial distribution and inheriting velocity from the local fluid flow.
In this study, the focus is on atmospheric clouds, where particles represent small water droplets (of radii ranging from $20$~$\upmu\text{m}$ to $60$~$\upmu\text{m}$) and the carrier fluid is atmospheric air.
Note, that for the remainder of the thesis terms ``particles'' and ``droplets'' shall be used interchangeably.
These particles confine simulations to specific physical dimensions and such conversion of spectral DNS units to physical units is based on matching the Kolmogorov scales.
In this study, we set: $\nu_{\text{air}} = 0.17 \, \text{cm}^{2} / \text{s} $ and $\epsilon_{\text{cloud}} = 400 \, \text{cm}^{2} / \text{s}^{3}$, which correspond to conditions observed in clouds with moderate to strong convection.
In these conditions the side length of domain under consideration turns out to be approximately 33 cm.

For particle-laden flows, in general, the statistics of droplets are mainly functions of three independent nondimensional parameters.
Firstly, the particle inertia can be expressed in terms of the Stokes number, $St = \tau_p / \tau_k$, defined as the ratio of the inertial response time to the
Kolmogorov time scale.
Furthermore, the range of scales of turbulence in the system which affects particle behaviour is measured by Taylor-microscale Reynolds number, $R_{\lambda}$, as defined in Section \ref{ssc:ch2.flow.bstat}, which is similar for all simulations under OWC (both DNS and LES) at approximately $165$.
Finally, when settling particles are considered, we use normalised terminal velocity, $S_V = V_T / v_k$, that is the particle still-fluid terminal velocity, $V_T$, normalised by the Kolmogorov velocity scale, $v_k = \eta / \tau_k$.
Note, that independence of these parameters remains true for simulations under one-way momentum coupling.
This picture, however, becomes more complex when two-way coupling is considered.
When effects of particles on the flow are taken into account, these parameters become indirectly dependent, e.g. large settling particles (with larger $St$ and $S_V$) are known to create additional vortical structures in flow, affecting $R_{\lambda}$, which---in turn---may alter settling velocity of particles (and, thus, $S_V$).

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrrrrrrrr}
$\mathbf{a}$ [$\upmu\text{m}$] & 20 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60  \\ \hline
$St$ & 0.254 & 0.396 & 0.571 & 0.777 & 1.015 & 1.284 & 1.585 & 1.918 & 2.283 \\
$S_V$ & 1.789 & 2.790 & 4.018 & 5.468 & 7.143 & 9.040 & 11.160 & 13.504 & 16.071 \\
$Fr$ & 0.81 & 3.08 & 9.21 & 23.23 & 51.76 & 104.93 & 197.44 & 349.77 & 589.54 \\
$\eta / a$ & 29.60 & 23.69 & 19.73 & 16.91 & 14.80 & 13.16 & 11.84 & 10.76 & 9.87 \\
$\Delta x_{\text{DNS}} / a$ & 64.58 & 51.67 & 43.06 & 36.90 & 32.29 & 28.70 & 25.83 & 23.48 & 21.53 \\
$\Delta x_{\text{LES}} / a$ & 219.62 & 175.70 & 146.42 & 125.50 & 109.81 & 97.61 & 87.85 & 79.86 & 73.21 \\
$\Delta t / \tau_p \times 10^2$ & 17.21 & 11.02 & 7.65 & 5.62 & 4.30 & 3.40 & 2.75 & 2.28 & 1.91 \\ \hline
\end{tabular}
\caption{Basic parameters for particle-laden flows for the range of particle radii (in $\upmu\text{m}$) that were used in this study.
$S_V$ and $Fr$ are only applicable to simulations with gravity.
Note, that values of $\Delta x / a$ ratio are listed separately for DNS and LES, since their spatial step $\Delta x$ differ significantly due to much smaller density of grid points in LES without changing domain size. }
\label{tab:part-stats}
\end{table}

Theses parameters are presented in Table \ref{tab:part-stats} for simulations in this study, relative to droplet radii.
In addition, the Froude number, $Fr$, is defined as the ratio of particle response time to the residence time of the particles in a Kolmogorov eddy.
Its value, however, may be expressed in terms of basic parameters introduced above, as $Fr = St \cdot S_V^2$.
Properties of general particle-laden flows are often dependent on density ratio of two phases, but due to focus of this study on atmospheric phenomena, such ratio remains constant, as
$$
\frac{\rho_p}{\rho} 
= \frac{\rho_{\text{water}}}{\rho_{\text{air}}} 
= \frac{1 \, \text{g} / \text{cm}^3}{0.001 \, \text{g} / \text{cm}^3} 
= 10^3 .
$$
For a broader picture, an extended analysis concerning the influence of particle–fluid density ratio on the dynamics of finite-size particles in homogeneous
isotropic turbulent flows can be found in \textcite{Shen2021}.
Since the density ratio is fixed in this study the particle mass loading, $\Phi_m$, is exclusively used as a measure of relative amount of particles in a system as it is directly related to the volume fraction by $\Phi_V = \Phi_m \cdot 10^{-3}$ (see Section \ref{ssc:ch2.flow.part}).
On that note, whenever necessary to obtain systems with larger mass loading, the super-particle parameterisation (where single computational particle represents several physical particles) was used.
For more information in which simulations super-particles were applied, see Appendix \ref{app:spp}.

Table \ref{tab:part-stats} also includes some spatial and temporal ratios used as control parameters.
Relatively large values of $\eta / a$ and $\Delta x / a$ indicate that indeed particles used in these simulations are very small when compared to both computational grid cells and smallest scales of turbulent eddies.
This second observation is paramount, as it not only makes the point-particle approximation justified.
It is also important for simulations under two-way momentum coupling, as it allows all computations of the interphase momentum transfer to use a simplified model that is valid under assumption that indeed $a \ll \eta$, making additional corrections not necessary \parencite{Horwitz2016}.

Due to the influence of momentum transfer from particles to the fluid some simulations under two-way momentum coupling required reduction of time step to preserve numerical stability of the simulation.
The base value used ($\Delta t = 9 \times 10^{-4}$) takes into account the presence of particles \parencite{Zhou2001} and is adjusted for interactions between particles (as modelled by their hydrodynamic interactions) by satisfying condition $\Delta t / \tau_p < 0.15$, where $\tau_p$ denotes particle response time, that was suggested by \textcite{Ayala2007}.
Only slight exceptions are simulations with the smallest droplets ($20$~$\upmu\text{m}$) where this ratio is $0.17$. 
Despite that, for some simulations $\Delta t$ needed to be further reduced even by a factor of $4$.
This was most common for simulations with settling particles and high super-particle factor, $M$, that often coincided with high particle mass loading.
These problems with stability appear to be caused by spikes in values of velocity field gradient that are inflated by momentum transfer from particles (i.e. $\mathbf{f}^{(p)}$ term in Equation \ref{eqn:n-s})
It locally saturates capabilities of the pseudo-spectral method and destabilises entire simulation (or, in other words, pushes ``effective'' CFL beyond required threshold of $0.3$).

These issues require separate, more systematic treatment, but based on the experimental data obtained in this study (i.e. observation of cases where $\Delta t$ required adjustment) some initial observations may be put forward.
First of all, the risk of such oversaturation is related to the average number of particles per grid node which depends on the total number of particles and their masses (radii), as well as simulation method that was used.
As outlined before, we employ LES to model turbulence with similar characteristics as by DNS with higher grid resolution.
Still, the domain for both methods is being populated with the same number of particles.
Hence, with grid sizes used here, it means that on average $64$ times more particles influence single node in LES than in DNS.
Moreover, this is exacerbated by preferential particle concentration (as measured by radial distribution function, see Section \ref{ssc:ch2.coll.rdfrrv}) that makes some nodes more saturated by particles than others.

In similar fashion, the super-particle parameterisation (i.e. when $M > 1$) may introduce non-physical clustering effect as single computational particle carries $M$ times more momentum.
This effect may be even more pronounced, since projection of particle momentum onto nearest neighbouring node depends on its distance to the position of that node.
For large values of $M$ much larger mass is concentrated at a single point which may happen to be located very close to an actual node and impart its huge momentum contribution, causing affected node to be oversaturated.
For smaller $M$ (ideally $M=1$) particle mass is expected to be more evenly distributed around node locations, thus making such adverse occurrences statistically much less likely.
